---
title: "005_AllBatches_Exploratory"
author: "Mike Martinez"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to redo the exploratory analysis but including all the samples from batch 1 as well. This code will be strucutred very similarly to `002_FlowData_Exploratory.Rmd`

---
title: "002_FlowData_Exploratory"
author: "Mike Martinez"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to use the outputs generated in `001_FlowData_Reformatted.Rmd` to generate paired paired boxplots
We need to also assess the normality of the data before we generate these boxplots so we can use the appropriate statistical test.

```{r set working directory and load libraries}

# Set working directory
setwd("/Users/mikemartinez/Desktop/Walnut_Project/Ongoing_Trial/PBMCs/Exploratory_Analysis/Batches_1_2_3/")

library(ggplot2)
library(ggpubr)
```

First, we need to specify the folder that holds all the files. We will use the path to the directory that holds all the files, as well as the file name to generate a full file path. We will use the file paths to store the data into two lists, one containing all the lymphocyte data, the other, all the monocyte data.
```{r specify all the lymphocyte and monocyte files}

# Set the path to the reformatted flow files
lymphocyte_folder <- "/users/mikemartinez/Desktop/Walnut_Project/Ongoing_Trial/PBMCs/Data/Reformatted_Data/Batches_1_2_3/Batch1_2_3_Lymphocytes/"
monocyte_folder <- "/users/mikemartinez/Desktop/Walnut_Project/Ongoing_Trial/PBMCs/Data/Reformatted_Data/Batches_1_2_3/Batch1_2_3_Monocytes/"

# List the files
lymphocyte_files <- list.files(lymphocyte_folder, pattern = ".csv$", full.names = FALSE)
monocyte_files <- list.files(monocyte_folder, pattern = ".csv$", full.names = FALSE)

# Initialize an empty list to hold all lymphocyte and monocyte files
lymphocyte_results <- list()
monocyte_results <- list()

# For loop to read in the file and store its contents in the lymphocyte_results list
for (file in lymphocyte_files) {
  
  # Define the full file path
  full_path <- file.path(lymphocyte_folder, file)

  # Read in the current file
  current_file <- read.csv(full_path, header = TRUE, sep = ",")
  print(dim(current_file))
  
  # Store the data frame in the list
  lymphocyte_results[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}

# For loop to read in the file and store its contents in the monocyte_results list
for (file in monocyte_files) {
  
  # Define the full file path
  full_path <- file.path(monocyte_folder, file)

  # Read in the current file
  current_file <- read.csv(full_path, header = TRUE, sep = ",")
  print(dim(current_file))
  
  # Store the data frame in the list
  monocyte_results[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}

```
Let's assess the normality so we can choose an appropriate test to use. To do this, we will use the Shapiro-Wilks test.
If the pvalue is < 0.05, you can reject the null hypothesis that the data is normal (i.e. the data is not normal)
```{r shapiro-wilks test}
# Iterate through the lymphocyte results and assess normality
for (result in names(lymphocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Run the Shapiro test
  shapiro_test_result <- shapiro.test(lymphocyte_results[[result]]$Statistic)
  print(result)
  print(shapiro_test_result)
}

# Iterate through the monocyte results and assess normality
for (result in names(monocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Run the Shapiro test
  shapiro_test_result <- shapiro.test(monocyte_results[[result]]$Statistic)
  print(result)
  print(shapiro_test_result)
 
}
```

Now that we have statistically assess the normality of the data with the Shapiro test, let's also visualize it graphically in the form of a histogram.
```{r histogram}

# Iterate through the lymphocyte results and assess normality
for (result in names(lymphocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Create a png plot
  png(paste(title, "histogram.png", sep = "_"))
  # Plot histogram of cells
  hist(lymphocyte_results[[result]]$Statistic, main = title, breaks = 10, xlab = "Number of Cells")
  dev.off()
}

# Iterate through the monocyte results and assess normality
for (result in names(monocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Create a png plot
  png(paste(title, "histogram.png", sep = "_"))
  # Plot histogram of cells
  hist(monocyte_results[[result]]$Statistic, main = title, breaks = 10, xlab = "Number of Cells")
  dev.off()
}
```

Since the majority of the data are not normal, we will use the Paired Wilcox test, which is non-parametric. The following loop will iterate through the files and generate the paired boxplots.
```{r plot}

# Plot for lymphocytes
for (result in names(lymphocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Plot
  pairedPlot <- ggpaired(lymphocyte_results[[result]], x = "Visit", y = "Statistic", id = "Sample_Number",
                  fill = "Visit", line.color = "black", line.size = 0.3) +
  labs(y = "Percentage",
       x = "Visit") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set3") +
    ggtitle(title) +
    theme(axis.text.x = element_text(size = 14),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_text(size = 14),
          axis.title.y = element_text(size = 16),
          title = element_text(face = "bold", size = 20)) +
  stat_compare_means(method = "wilcox.test", paired = TRUE, vjust = 1.2)
ggsave(pairedPlot, file = paste(title, "Paired_Plot.png", sep = "_"))
}

# Plot for monocytes
for (result in names(monocyte_results)) {
  
  # Make the result name a little nicer
  title <- sub("\\.csv$", "", result)
  title <- sub("Merged_2\\.3_", "", title)
  
  # Plot
  pairedPlot <- ggpaired(monocyte_results[[result]], x = "Visit", y = "Statistic", id = "Sample_Number",
                  fill = "Visit", line.color = "black", line.size = 0.3) +
  labs(y = "Percentage",
       x = "Visit") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set3") +
    ggtitle(title) +
    theme(axis.text.x = element_text(size = 14),
          axis.title.x = element_text(size = 16),
          axis.text.y = element_text(size = 14),
          axis.title.y = element_text(size = 16),
          title = element_text(face = "bold", size = 20)) +
  stat_compare_means(method = "wilcox.test", paired = TRUE, vjust = 1.2)
ggsave(pairedPlot, file = paste(title, "Paired_Plot.png", sep = "_"))
}
```













