---
title: "001_FlowData_Reformatted.Rmd"
author: "Mike Martinez"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to reformat the original flow cytometry data from batch 2 and 3 from the walnut PBMC project. This is to combine both data files into one master data file, encompassing 32 samples (16 before and 16 after.) 

```{r set working directory and load libraries}
setwd("/users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Reformatted_Data/")

library(tidyverse)
library(dplyr)
```

First, we will read in both data files (batch 2 and batch 3).
Add a column for batch number for each dataframe.
Concatenate them into one dataframe
```{r read in the data files}
b2 <- read.csv("/Users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Original_Data_From_Core/Flow_Batch_2.csv", header = TRUE, sep = ",")
b2$Batch <- "2"
b3 <- read.csv("/Users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Original_Data_From_Core/Flow_Batch_3.csv", header = TRUE, sep = ",")
b3$Batch <- "3"
flow <- rbind(b2, b3)
```

Let's begin some initial processing of the dataframes.
First, we do not need the observations for fcs or scatter. We need everything starting from the live cells.
```{r remove observations for first 2 depth levels}
# Depth "0" indicates the overall .fcs
# Depth "> " indicates the scatter. NOTE: THERE IS WHITE SPACE AFTER THE ">"
# For instances where the depth is 0 or >, we can remove these rows

flow <- flow[!(flow$Depth == "0" | flow$Depth == "> "),]
```

Now, I want to make the data frame a little more easily parseable. 
```{r}
setwd("/Users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Reformatted_Data/")
# Separate out the name column in Well_Number, Sample_Number, Visit, and Rest
flow1 <- flow %>%
  separate(Name, into = c("Well_Number", "Sample_Number", "Col2"), sep = " ", extra = "merge") 
  
# Add a time column telling whether or not the sample is before or after walnut
Before <- "V2"
flow1$Visit <- ifelse(grepl(Before, flow1$Col2), "Before", "After")

# We can take the live single cell data and store it in another data frame
liveSC <- flow1[flow1$Col2 == "V2.fcs/scatter/live/Single Cells" | flow1$Col2 == "V3.fcs/scatter/live/Single Cells",]
flow1 <- flow1[!(flow1$Col2 == "V2.fcs/scatter/live/Single Cells" | flow1$Col2 == "V3.fcs/scatter/live/Single Cells"),]
write.csv(liveSC, file = "Merged_2.3_All_Live_SingleCells.csv")
write.csv(flow1, file = "Merged_2.3_All_Flow_Data.csv")
```

We can isolate just the lymphocytes and their subcategories into separate data frames
```{r isolate lymphocyte markers}

setwd("/Users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Reformatted_Data/Batch2_3_Lymphocytes/")

# All lymphocyte data
All_Lymphocytes <- flow1[grepl("Lymphocytes", flow1$Col2),]
write.csv(All_Lymphocytes, file = "Merged_2.3_All_Lymphocytes.csv")

Overall_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes",]
write.csv(Overall_Lymphocytes, file = "Merged_2.3_Overall_Lymphocytes.csv")

CD3_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD3" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD3",]
write.csv(CD3_Lymphocytes, file = "Merged_2.3_CD3_Lymphocytes.csv")

CD4_pos_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q1: CD8-PerCP-Cy5.5-A- , CD4-PE (R-phycoerythrin)-A+" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q1: CD8-PerCP-Cy5.5-A- , CD4-PE (R-phycoerythrin)-A+",]
write.csv(CD4_pos_Lymphocytes, file = "Merged_2.3_CD4_pos_CD8_neg_Lymphocytes.csv")

Double_pos_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q2: CD8-PerCP-Cy5.5-A+ , CD4-PE (R-phycoerythrin)-A+" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q2: CD8-PerCP-Cy5.5-A+ , CD4-PE (R-phycoerythrin)-A+",]
write.csv(Double_pos_Lymphocytes, file = "Merged_2.3_CD4_CD8_Double_pos_Lymphocytes.csv")

CD8_pos_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q3: CD8-PerCP-Cy5.5-A+ , CD4-PE (R-phycoerythrin)-A-" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q3: CD8-PerCP-Cy5.5-A+ , CD4-PE (R-phycoerythrin)-A-",]
write.csv(CD8_pos_Lymphocytes, file = "Merged_2.3_CD8_pos_CD4_neg_Lymphocytes.csv")

Double_neg_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q4: CD8-PerCP-Cy5.5-A- , CD4-PE (R-phycoerythrin)-A-" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD3/Q4: CD8-PerCP-Cy5.5-A- , CD4-PE (R-phycoerythrin)-A-",]
write.csv(Double_neg_Lymphocytes, file = "Merged_2.3_CD4_CD8_Double_neg_Lymphocytes.csv")

CD19_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/CD19" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/CD19",]
write.csv(CD19_Lymphocytes, "Merged_2.3_CD19_Lymphocytes.csv")

NK_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/NK" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/NK",]
write.csv(NK_Lymphocytes, "Merged_2.3_NK_Lymphocytes.csv")

NKT_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/NKT" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/NKT",]
write.csv(NKT_Lymphocytes, "Merged_2.3_NKT_Lymphocytes.csv")

CD16_CD3_doubleNeg <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/Q4: CD16-PE-Cy7-A- , CD3-APC-H7-A-" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/Q4: CD16-PE-Cy7-A- , CD3-APC-H7-A-",]
write.csv(CD16_CD3_doubleNeg, "Merged_2.3_CD16_CD3_Double_neg_Lymphocytes.csv")

T_Lymphocytes <- All_Lymphocytes[All_Lymphocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Lymphocytes/T" | All_Lymphocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Lymphocytes/T",]
write.csv(T_Lymphocytes, "Merged_2.3_T_Lymphocytes.csv")

```

Now we can do the same thing for the monocytes
```{r isolate monocyte markers}

setwd("/Users/mikemartinez/Desktop/Walnut_Project/PBMCs/Data/Reformatted_Data/Batch2_3_Monocytes/")

# All monocytes data
All_Monocytes <- flow1[grepl("Monocytes", flow1$Col2),]
write.csv(All_Monocytes, file = "Merged_2.3_All_Monocytes.csv")

Overall_Mono <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes",]
write.csv(Overall_Mono, file = "Merged_2.3_Overall_Monocytes.csv")

CD64_Mono <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes/CD64+" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes/CD64+",]
write.csv(CD64_Mono, file = "Merged_2.3_CD64_Monocytes.csv")

Classical_Mono <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes/CD64+/classical mono" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes/CD64+/classical mono",]
write.csv(Classical_Mono, file = "Merged_2.3_Classical_Monocytes.csv")

int_Mono <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes/CD64+/int mono" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes/CD64+/int mono",]
write.csv(int_Mono, file = "Merged_2.3_Int_Monocytes.csv")

NC_Mono <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes/CD64+/nonclassical mono" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes/CD64+/nonclassical mono",]
write.csv(NC_Mono, file = "Merged_2.3_Nonclassical_Monocytes.csv")

CD14_CD16_doubleNeg <- All_Monocytes[All_Monocytes$Col2 == "V2.fcs/scatter/live/Single Cells/Monocytes/CD64+/Q4: CD14-FITC-A- , CD16-PE-Cy7-A-" | All_Monocytes$Col2 == "V3.fcs/scatter/live/Single Cells/Monocytes/CD64+/Q4: CD14-FITC-A- , CD16-PE-Cy7-A-",]
write.csv(CD14_CD16_doubleNeg, file = "Merged_2.3_CD14_CD16_Double_Neg_Monocytes.csv")
```



